## Overview: SQL and R Assignment - Store survey data in SQL database and then load into an R dataframe
Surveyed 7 friends to rate the following 6 recent movies on a scale of 1 (Hated it) to 5 (Absolutely Loved it):
1) Top Gun - Maverick
2) Dr. Strange Multiverse of Madness
3) Thor: Love & Thunder
4) Lightyear
5) Bullet Train
6) Jurassic World: Dominion

The results (observations) were inserted into a MySQL database table D607_Wk3_Survey created with the following SQL Script
## Insert survey observations into MySQL database
CREATE TABLE D607_Wk3_Survey
(
 participant int not null,
  top_gun int NULL,
  dr_strange int NULL,
  thor  int NULL,
  lightyear  int NULL,
  bullet_train  int NULL,
  jurassic_world  int NULL
);

INSERT INTO D607_Wk3_Survey VALUES (1,4,4,3,3,5,3);
INSERT INTO D607_Wk3_Survey VALUES (2,3,null,null,3,4,null);
INSERT INTO D607_Wk3_Survey VALUES (3,null,5,4,4,5,3);
INSERT INTO D607_Wk3_Survey VALUES (4,null,4,3,null,null,null);
INSERT INTO D607_Wk3_Survey VALUES (5,null,null,null,null,5,null);
INSERT INTO D607_Wk3_Survey VALUES (6,null,4,null,4,null,null);
INSERT INTO D607_Wk3_Survey VALUES (7,null,null,4,3,null,null);

```{r}
library(RMySQL)
```
## Import Data:
```{r}
mydb = dbConnect(RMySQL::MySQL(), dbname='tb', user='root', password='password', host='localhost',port=3306)

# Save sql table into dataframe called movies
movies <- dbGetQuery(mydb, "select * from D607_Wk3_Survey")
```
## Null Values:
```{r}

str(movies)
```
## Conclusion:
Selected 6 recent movies and surveyed 7 participants. If the participant did not watch the movie as of 9/11 then value was left null. Created a MySQL database with table D607_Wk3_Survey, inserted data points and then imported into R. Checked how null values were treated and values are now NA.
