---
title: "607 Wk 4 Assignment"
output: html_document
date: "2022-09-18"
---
## Overview: Data 607 Week 4 Assignment covers R Character Manipulation

## 1) Identify Data and Statistics Majors
#Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"
# Import Data:
```{r}
library(tidyverse)

url = 'https://raw.githubusercontent.com/fivethirtyeight/data/2d2ff3e9457549d51f8e571c52099bfe9b2017ad/college-majors/majors-list.csv'
collmajors <- read.csv(url, header = TRUE, sep = ",", stringsAsFactors = FALSE)
stat_data <- grep(pattern = 'STATISTICS|DATA', collmajors$Major, value = TRUE, ignore.case = TRUE)
stat_data
```
# Alternatively, could also identify values by: 
```{r}
majors <- as.character(collmajors$Major)
str_extract(majors,"STATISTICS|DATA")
```
# List the majors based on results of str_extract
```{r}
print(majors[44])
print(majors[52])
print(majors[59])
```

## 2) Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
#Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
```{r}
##Save text input as fruits
fruits <- '
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
'
```

```{r}
fruits_2 <- str_extract_all(fruits, "\\w[a-z]+\\s?[a-z]+\\w")
dput(fruits_2)
```
```{r}
# Fruits_2 is classed as a list which is why it's showing up in the output so need to unlist
fruits_3 <- unlist(fruits_2)
dput(fruits_3)
```

## 3) Describe, in words, what these expressions will match:
```{r}
# Set up Example to be used to show matches for each expression
tst <- c("aaa", "abba", "abab", "abaca","abcFILLERcba")
```

(.)\1\1 - Any character repeated 3 times consecutively. For testing example, will need to input in quotes and add additional backslashes
```{r}
str_extract(tst,"(.)\\1\\1")
```
"(.)(.)\\2\\1" - A pair of characters followed by the same pair of characters in reversed order
```{r}
str_extract(tst,"(.)(.)\\2\\1")
```
(..)\1 - Any two characters repeated.  For testing example, will need to input in quotes and add additional backslash
```{r}
str_extract(tst,"(..)\\1")
```
"(.).\\1.\\1" - A character followed by any character, the original character, any other character, the original character again.
```{r}
str_extract(tst,"(.).\\1.\\1")
```
"(.)(.)(.).*\\3\\2\\1" -Three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order
```{r}
str_extract(tst,"(.)(.)(.).*\\3\\2\\1")
```

## 4) Construct regular expressions to match words that:
```{r}
# Set up Example to be used to show matches for each expression
tst4 <- c("alexa", "church", "eleven")
```
Start and end with the same character
#    "^(.).*\\1$"
```{r}
str_extract(tst4,"^(.).*\\1$")
```
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.) 
#   "([a-zA-Z][a-zA-Z]).*\\1"
```{r}
str_extract(tst4,"([a-zA-Z][a-zA-Z]).*\\1")
```
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
#   "([a-zA-Z]).*\\1.*\\1.*"
```{r}
str_extract(tst4,"([a-zA-Z]).*\\1.*\\1.*")

```

