---
title: "607 Wk 7 Project 2 - Gross Domestic Product Analysis"
author: "Jennifer Abinette"
output: html_document
date: "2022-09-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```
### Introduction: This analysis looks at trends in National Gross Domestic Products over the last 10 quarters.

## 1) Download Gross Domestic Products dataset from https://apps.bea.gov/iTable/iTable.cfm?reqid=19&step=3&isuri=1&nipa_table_list=5&categories=survey

## 2) Save a copy as CSV file

## 3) Load CSV file saved in Github
```{r}
url = 'https://raw.githubusercontent.com/JAbinette/CUNY-607-DataAcquisition/main/Project%202%20-Gross%20Domestic%20Product%2009-29-22.csv'
d.load1 <- read.csv(url, header = TRUE, sep = ",", skip=4, stringsAsFactors = FALSE)

head(d.load1)
```
## 4) Tidy and Transform Data for Analyses

```{r}
# Rename Columns
names(d.load1) <- c("Line", "Description", "March 2020", "June 2020", "September 2020", "December 2020", "March 2021", "June 2021", "September 2021", "December 2021", "March 2022", "June 2022")

# Remove 1st Record as moved Quarter information into column name above
GDP1 <- subset(d.load1[-1,])

# Trim Whitespaces in Description
GDP1$Description <- trimws(GDP1$Description, whitespace = "[ \t\r\n]")

# Adjust Quarters to numeric data type rather than character
GDP1$'March 2020' <- as.numeric(GDP1$'March 2020')
GDP1$'June 2020' <- as.numeric(GDP1$'June 2020')
GDP1$'September 2020' <- as.numeric(GDP1$'September 2020')
GDP1$'December 2020' <- as.numeric(GDP1$'December 2020')
GDP1$'March 2021' <- as.numeric(GDP1$'March 2021')
GDP1$'June 2021' <- as.numeric(GDP1$'June 2021')
GDP1$'September 2021' <- as.numeric(GDP1$'September 2021')
GDP1$'December 2021' <- as.numeric(GDP1$'December 2021')
GDP1$'March 2022' <- as.numeric(GDP1$'March 2022')
GDP1$'June 2022' <- as.numeric(GDP1$'June 2022')

head(GDP1)
```
## 5) Analyze Overall Trend for GDP 
```{r}
# Subset data to only Gross Domestic Product
GDP_Only <- subset(GDP1, Line == '1')
head(GDP_Only)
```

## Conclusion: The analysis of the overall trend of Gross Domestic Product is that in almost all cases (except June 2020), GDP has increased from the previous quarter.  As shown above, at the end of Quarter 1, 2020 (March 2020) GDP was at $21,538 billion comparted to Quarter 2, 2022 (June 2022) at $25,248.5.  If GDP is actually an accurate representation of the economic health of a country then an increase is a great sign.
