---
title: "607 Wk 7 Project 2 - New York City Typical Home Vale Analysis"
author: "Jennifer Abinette"
output: html_document
date: "2022-09-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```
## 1) Download dataset from https://www.zillow.com/research/data/
Home Values Section: 
Data Type = ZHVI Single-Family Homes Time Series;
Geography = Zip Code

## 2) Remove records where State is not NY from the CSV file as file is too large for Github and analysis is on NYC

## 3) Load CSV file saved in Github
```{r}
url = 'https://raw.githubusercontent.com/JAbinette/CUNY-607-DataAcquisition/main/Project%202%20-Zillow%20NY%20Only.csv'
d.load <- read.csv(url, header = TRUE, sep = ",", stringsAsFactors = FALSE)
str(d.load)
```
## 4) Subset data to include only NYC records and Remove extraneous columns
```{r}
nyc <- subset(d.load, City=='New York', select = c(City, X8.31.2021, X8.31.2022) )

head(nyc)
```
## 5) Transform Data
```{r}
# Rename Columns
names(nyc) <- c("City", "August_2021", "August_2022")

# Create variables finding Average for all of New York City 
nyc_1 <- nyc  %>% 
  group_by(City) %>%
  mutate (Avg_2021 = mean(August_2021, na.rm = TRUE)) %>%
  mutate (Avg_2022 = mean(August_2022, na.rm = TRUE)) 

# Create new data frame removing extraneous columns and duplicate rows
nyc_2 <- as.data.frame( distinct( subset (nyc_1, select = c (City, Avg_2021, Avg_2022) ) ) )
head(nyc_2)
```
## 6) Analyze Percent Change between 2021 and 2022
```{r}
# Add Percent Growth Variable
nyc_2$Growth = 100 * ( (nyc_2$Avg_2022 - nyc_2$Avg_2021) / nyc_2$Avg_2021 )
print(nyc_2)
```
## Conclusion: The analysis indicates that there was a 7% growth in typical home value for New York, NY where the average increased from $1,317,949 in August 2021 to $1,409,759 in August 2022.